---
title: "Data Mining: Final Project"
author: "David Fraire"
date: "5/16/2019"
output: pdf_document
---

```{r Library & Data Loading, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Code for loading Library Packages
library(matrixStats)
library(readr)
library(dbplyr)
library(dplyr)
library(ggformula)
library(ggplot2)
library(ggstance)
library(grDevices)
library(haven)
library(markdown)
library(Matrix)
library(pander)
library(stargazer)
library(stringr)
library(tidyr)
library(tidyverse)

final <- read.csv("../data/finaldataset.csv", header=TRUE)

```
## Effects of Head Start on PPVT at Age 3, College enrollment, & Highschool Graduation
The following regressions show in table form will represent the effects of the Headstart program on the outcome variables of **PPVTat3**, **somecollege**, **hsgrad**. We decided to use a mixture of linear models and logit models to proceed with analysis of the effect of Headstart program on the three aforementioned dependent variables. The table showing effects of Head Start on **PPVTat3** is a linear model. The tables showing Head Start effects on **somecollege** and **hsgrad** are logit models and the shown coefficients have been exponentiated for interpretation reasons. 

``` {r Models, echo=FALSE, results='asis'}
## Linear Models for PPVTat3
lmPPVT0 = lm(PPVTat3~headstart, data=final)
lmPPVT1 = lm(PPVTat3~headstart + Male , data=final)
lmPPVT2 = lm(PPVTat3~headstart + Hispanic + Black + Male, data=final)
lmPPVT3 = lm(PPVTat3~headstart + Hispanic + Black + Male + hsgrad + FirstBorn + headstart*Black + headstart*Hispanic, data=final)


## Logit models for College + Odds Ratios
glmCE0<- glm(somecollege~headstart, data=final, family=binomial)
glmCE1<- glm(somecollege~headstart + Male, data=final, family=binomial)
glmCE2<- glm(somecollege~headstart + Male + Black + Hispanic, data=final, family=binomial)
glmCE3<- glm(somecollege~headstart + Male + Black + Hispanic + LogInc_0to3+ headstart*Black + headstart*Hispanic, data=final, family=binomial)


Odd0<- exp(coef(glmCE0))
Odd1<- exp(coef(glmCE1))
Odd2<- exp(coef(glmCE2))
Odd3<- exp(coef(glmCE3))


## Logit models for highschool grad (highschool is = AGEHSGRAD<12)
glmhs0<- glm(hsgrad~headstart, data=final, family=binomial)
glmhs1<- glm(hsgrad~headstart + Male , data=final, family=binomial)
glmhs2<- glm(hsgrad~headstart + Male + Black + Hispanic , data=final, family=binomial)
glmhs3<- glm(hsgrad~headstart + Male + Black + Hispanic +LogInc_0to3 + MothED + headstart*Black + headstart*Hispanic, data=final, family=binomial)


OddHS0<- exp(coef(glmhs0))
OddHS1<- exp(coef(glmhs1))
OddHS2<- exp(coef(glmhs2))
OddHS3<- exp(coef(glmhs3))
```

``` {r PPVT Table, echo=FALSE, results='asis'}

stargazer::stargazer(lmPPVT0, lmPPVT1,lmPPVT2, lmPPVT3,type="latex", title="Headstart effect on PPVT Scores at Age 3",header=FALSE,single.row=TRUE, column.sep.width = ".6pt", omit.stat = c("f"))

```
For the previous table, we begin in the first by regressing the data for **PPVTat3** only on **headstart** using a linear fit. The effect of the headstart program in this model shows a very statistically significant negative effect of -6.741. The linear model in the last column includes **Hispanic**, **Black**, **Male**, **hsgrad**, **FirstBorn**, and interaction terms for **headstart**. The very significant variables from this regression are **headstart**, **Hispanic**, **Black**, **Male**, **hsgrad**, and **FirstBorn**. The **Black** variable has a very large average effect on the outcome of PPVT scores and is the highest in magnitude. The **headstart** variable still has a significant effect but is lower in magnitude and lower in signiicance than that of previous regressions with less covariates. This means that participation in Headstart has a consistently negative effect on the test scores with the data provided in our dataset even after controlling for other highly significant variables. Given that the children who participate in Headstart are likely to come from disadvatanged socioeconomic backgrounds, this can be explained in the following models as those who are in Headstart will likely have lower scores to start. The last model included variables that were shown to explain a significant amount of variation in the original dataset and were therfore chosen to be included in order to convey the decreasing marginal effect of Headstart on PPVT scores at age 3. 

``` {r College Table, echo=FALSE, results='asis'}
stargazer::stargazer(glmCE0, glmCE1, glmCE2, glmCE3, type="latex", title="Headstart effect on College Enrollment",header=FALSE,single.row=TRUE,apply.coef = exp)
```

For the effect of Head Start on college enrollment which is measured by the binary variable **somecollege**, we used a logit model to see the effects on the odds of having gone to college. Note, that the coefficients in the table have been exponentiated in order to improve tr Simply running a regression of **headstart** on **somecollege** shows that **headstart** has a significant effect on the outcome of having enrolled at a college. The odds of attending college are increased by a factor of 1.625. However, this effect is shown to decrease and even become statistically insignificant when controlling for other significant variables such as **Male**, **Black**, **Hispanic**, and **LogInc_0to3**. 

```{undefined eval=FALSE, include=FALSE}
stargazer::stargazer(Odd0,Odd1,Odd2,Odd3, type="latex",header=FALSE,single.row=TRUE, title="Log Odds of Variable", flip = TRUE)
```
The following tables tell us the log odds of each of the regressions showing that **headstart** in the regression with all included covariates has a positive effect of a factor of 1.059 but is not a statistically significant variable after controlling for **Male**, **Black**, **Hispanic**,**LogInc_0to3** and **headstart** interaction terms, all of which are very statistically significant. 

``` {r Highschool Table, echo=FALSE, results='asis'}
stargazer::stargazer(glmhs0, glmhs1, glmhs2, glmhs3, type="latex", title="Headstart effects on High School Graduation",header=FALSE,single.row=TRUE, apply.coef = exp)
```

In order to measure the effect of Head Start on a child graduating high school we decided to use a logit model in order to tell us the odds ratio. To start, we regressed **headstart** on **somecollege** and then exponentiated the coefficient for the all the regressions included in the table in order to tell us the effect on the odds of receiving the outcome of having gone to college for particular children. When regressing **headstart** only on **somecollege** we see that being enrolled in the program increases the odds of going to college by 1.625 and is very significant. Our last column includes the variables **Male**, **Male**,**Male**,**Male**,**Male**

The model uses a logit link funciton which tells us in column that the effect of particpating in the Headstart program will increase the log odds of graduating highschool by .486. 

```{undefined eval=FALSE, include=FALSE}
stargazer::stargazer(OddHS0, type="latex",header=FALSE,single.row=TRUE, 


```