---
title: "Davids_R_Script"
author: "David Fraire"
date: "5/16/2019"
output: pdf_document
---
Code for loading Library Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(foreign)
library(matrixStats)
library(readr)
library(FNN)
library(tidyverse)
library(datasets)
library(dbplyr)
library(dplyr)
library(ggformula)
library(ggplot2)
library(ggstance)
library(graphics)
library(grDevices)
library(haven)
library(lattice)
library(markdown)
library(Matrix)
library(methods)
library(mosaic)
library(mosaicData)
library(pander)
library(RColorBrewer)
library(stargazer)
library(stringr)
library(tidyr)
library(tidyverse)

```


The following regressions show in table form will represent the effects of the Headstart program on PPVTat3, College Education, and High School graduation. 

We decided to use a Genearlized Linear Models and a Linear Models to proceed with analysis of the effect of Headstart program. For the first model, we regressed the data for PPVTat3 on headstart using a linear fit. The effect of the headstart program in this model shows a negative 


#HS on PPVT Score
``` {r ,echo=FALSE, results='asis'}
final <- read.csv("../data/finaldataset.csv", header=TRUE)

lmPPVT0 = lm(PPVTat3~headstart, data=final)
lmPPVT1 = lm(PPVTat3~headstart + BirthWeight + Male , data=final)
lmPPVT2 = lm(PPVTat3~headstart + Hispanic + Black + Male, data=final)
lmPPVT3 = lm(PPVTat3~headstart + Hispanic + Black + Male + BirthWeight + hsgrad+ FirstBorn, data=final)

stargazer::stargazer(lmPPVT0, lmPPVT1, lmPPVT2, lmPPVT3 ,type="text")

```


#HS on College Enrollment
``` {r, echo=FALSE}

glmCE0<- glm(somecollege~headstart, data=final, family=binomial)
glmCE1<- glm(somecollege~headstart + Male, data=final, family=binomial)
glmCE2<- glm(somecollege~headstart + Male + Black + Hispanic, data=final, family=binomial)
glmCE3<- glm(somecollege~headstart + Male + Black + Hispanic + BirthWeight + LogInc_0to3, data=final, family=binomial)

stargazer::stargazer(glmCE0, glmCE1, glmCE2, glmCE3, type="text", title="Regressions of HS on College Enrollment")



```


#HS on High School grad
``` {r, echo=FALSE}
glmhs0<- glm(hsgrad~headstart, data=final, family=binomial)
glmhs1<- glm(hsgrad~headstart + Male , data=final, family=binomial)
glmhs2<- glm(hsgrad~headstart + Male + Black + Hispanic , data=final, family=binomial)
glmhs3<- glm(hsgrad~headstart + Male + Black + Hispanic +LogInc_0to3 + MothED, data=final, family=binomial)

stargazer::stargazer(glmhs0, glmhs1, glmhs2, glmhs3, type="text", title="Regressions of HS on High School Graduation")

```


