---
title: 'Short and Long Term Effects of Headstart'
author: "Frank Chou, David Fraire, Tejaswi Pukkalla"
date: "May 16, 2019"
output: pdf_document
---

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readstata13)
library(clusterSEs)
library(stargazer)
library(effects)
library(knitr)
```

# Abstract

# Introduction

Head Start is a nationwide preschool program targeted at children who come from socio-economically disadvantaged backgrounds. The program, started in 1965 as part of the 'War on Poverty', is funded federally. Head Start has helped over 30 million children and their families in all 50 states, the District of Columbia, Puerto Rico and the U.S. territories. It provides comprehensive early childhood education, health, nutrition, and parent involvement services to these children and their families. It helps create healthy development in children ages three to five from low-income families. 

The children are eligible based on the net family's income. Although each local program includes other eligibility criteria, such as disabilities and services needed by other family members. Families must earn less than the federal poverty level. Families sometimes may qualify under a categorical eligibility category. Kids from higher income families or families experiencing emergency situations can be enrolled up to 10% of any funded program. While the Obama administration consistently increased the funding for the Head Start program, the Trump administration has cut down the budget allocated for education and training outlays by about 45.7% - which severely impacted Head Start programs. This brought us to our research question: Is Head Start effective? Does it have any short-term and/or long-term benefit on its participants? Does this effect vary for different races? Many studies have been conducted during Head Start's multi-decade history to measure its effectiveness. Based on our literature review of the subject, there seems to be no academic or political consensus about the program's effectiveness

CCR Analytics published their results in 2014, studying 49,467 children assessed in the 2012-2013 school year from 81 Head Start programs throughout California. The study was open to all California Head Start programs participants who used the DRDP-PS 2010 assessment tool. It found that providing two years of Head Start to a child increases the probability by at least 13% that the child will be ready for school. In 2009, David Deming evaluated the program, using the National Longitudinal Survey of Youth (NLSY 79 data). He compared siblings and found that those who attended Head Start showed stronger academic performance, more likely to graduate from high school and attend college, and less likely to suffer from poor health as an adult. We are using Deming's dataset as our primary dataset to explore the causality effects and see if we can get similar results or have something new waiting for us.

According to Lee, who collected data across sixty Head Start classrooms in 2007 and 2008, Head Start is associated with significant gains in test scores and significantly reduces the probability that a child will repeat a grade. In 2005, Barnett and Hustedt reviewed the literature and stated that although the reviews found were mixed, they were generally positive. According to the Administrative History of the Office of Economic Opportunity, children who finish the program and are sent to disadvantaged schools perform worse than their peers by as early as second grade. Only by isolating such children and sending them to better schools could gains be sustained. Time magazine's columnist Joe Klein, in 2011, called for the elimination of Head Start, citing an internal undisclosed report that the program is costly and makes a negligible impact on children well-being. W. Steven Barnett, director of the National Institute for Early Education Research at Rutgers University, rebutted Klein stating that the weighing all of the evidence and not just that cited by partisans on one side or the other leads to the most accurate conclusion that Head Start produces modest benefits including some long-term gains for children.

# Methodology

Given that the National Longitudinal Survey of Youth consisted of survey participants from 1979 to 2004, there was a number of significant data processing steps needed prior to any analysis. With 12,686 participants spanning nearly two decades, the principal population we would be targeting their early childhood to young adulthood time frames. Individuals were interviewed annually until 1994, and then biennially  afterwards. The age of the cohort within the survey consisted of individuals born between 1957 and 1964. The gender ratio was split approximately 50:50 between men and women (6,403) males and (6,283) females. Our goal on understanding on the short and long term effects of the Head Start program, focused on the subset of children born to the female NLSY79 respondents. 

Using a subset of data created by Deming in his paper *"Early Childhood Intervention and Life-Cycle Skill Development: Evidence from Head Start,"* we utilized a subset of 11,470 children from the NLSY79 survey to conduct our paper. The first step was the identify which independent and dependent variables that we would examine. With variables ranging from whether or not a child had a disability to asthma, income levels at 3 years hold, to test scores on the Peabody Individual Achievement Test (PIAT) for both Math (PIATMT) and Reading Comprehension (PIAT), the goal was to read though the 882 variable list and find the features that we think would be useful to define both our research question, and provide the underlying dataset to demonstrate our findings. A summary of children who participated within the survey and are included in our research data set are included in the results section. With this in mind, we focused on the following variables to help us answer our question:

* **Composite Scores from ages 6 to 11**
* **Composite Scores from ages 12 to 14**
* **Composite Scores from ages 15 to 18**
* **Peabody Picture Vocabulary Test (PPVT) at 3 years old**
* **High School Graduation**
* **College Enrollment**
* **Grade Repetition**
* **Self Reported Health**

The **Composite Scores** were generated by selecting the number of children throughout the survey time frame who were at a given age at any given time. The composite score was generated from multiplying the percentile scores of each child from the Peabody Picture Vocabulary Test, Peabody Individual Achievement Math Test, Peabody Individual Achievement Reading Comprehension Test. The goal behind this design was such that all participating children within the study can be ranked and compared within each year's cohorts, as well as across time between cohorts. 

The **Peabody Picture Vocabulary Test** itself was the direct raw score from when the child was tested at 3 years old from the years 1986, 1988, and 1990. 

The **High School Graduation** was a indicator variable generated from whether or not a child every graduated from high school. Given that we have different cohorts of children each reaching and completing high school at different years, this indicator variable would be 0 for children who have not reached and completed a high school education, i.e. a child born in 1998, was aged 6 in the year 2004, and therefore, has a High School Graduation variable as zero.

**College Enrollment** was an indicator variable structured much like High School Graduation. If a child ever reached 13 years of educations, namely, a university/college institution, this variable would become 1 and 0 otherwise. Children who were too young, would still have an result of 0.

**Grade Repetition** was an indicator variable based on whether or not a child ever repeated one or more grades from primary (elementary), secondary (middle and high), and tertiary school (university) levels.

**Self Reported Health** was an indicator variable based on whether or not a child ever reported a less than 3 on a scale of 5 during the survey.

# Results and Analysis
```{r frank preprocessing, include=FALSE}
rm(list=ls())

library(tidyverse)
library(readstata13)
library(clusterSEs)
library(stargazer)
library(effects)
library(knitr)

df <- read.csv("../data/finaldataset.csv")
df2 <- read.dta13("../data/deming_final_data.dta")

# Preprocessing (some more)

df$hs.age <- rowMeans(df[,c("Age_1stHS88","Age_1stHS90","Age_1stHS92","Age_1stHS94","Age_1stHS96","Age_1stHS98",
                         "Age_1stHS100","Age_1stHS102","Age_1stHS98","Age_1stHS104")],
                   na.rm=TRUE)

calc_age <- function(currentyear,childyear){
  currentyear-childyear
}

df$childage1986 <- calc_age(1986,df2$DOB_Yr_Child)
df$childage1988 <- calc_age(1988,df2$DOB_Yr_Child)
df$childage1990 <- calc_age(1990,df2$DOB_Yr_Child)
df$childage1992 <- calc_age(1992,df2$DOB_Yr_Child)
df$childage1994 <- calc_age(1994,df2$DOB_Yr_Child)
df$childage1996 <- calc_age(1996,df2$DOB_Yr_Child)
df$childage1998 <- calc_age(1998,df2$DOB_Yr_Child)
df$childage2000 <- calc_age(2000,df2$DOB_Yr_Child)
df$childage2002 <- calc_age(2002,df2$DOB_Yr_Child)
df$childage2004 <- calc_age(2004,df2$DOB_Yr_Child)

## PIAT Participation

part.number <- rbind(
  11470-sum(is.na(df$compscoreage6)),
  11470-sum(is.na(df$compscoreage7)),
  11470-sum(is.na(df$compscoreage8)),
  11470-sum(is.na(df$compscoreage9)),
  11470-sum(is.na(df$compscoreage10)),
  11470-sum(is.na(df$compscoreage11)),
  11470-sum(is.na(df$compscoreage12)),
  11470-sum(is.na(df$compscoreage13)),
  11470-sum(is.na(df$compscoreage14)),
  11470-sum(is.na(df$compscoreage15)),
  11470-sum(is.na(df$compscoreage16)),
  11470-sum(is.na(df$compscoreage17)),
  11470-sum(is.na(df$compscoreage18))
)

## Age

part.age <- c(
  "6 Years Old",
  "7 Years Old",
  "8 Years Old",
  "9 Years Old",
  "10 Years Old",
  "11 Years Old",
  "12 Years Old",
  "13 Years Old",
  "14 Years Old",
  "15 Years Old",
  "16 Years Old",
  "17 Years Old",
  "18 Years Old"
)

## Counting by age

part_function.total <- function(age){
  part.df <<- subset(df,childage1986==age |childage1988==age |childage1990==age |childage1992==age |childage1994==age |
                       childage1996==age |childage1998==age |childage2000==age |childage2002==age |childage2004==age)
  nrow(part.df)
}

part.total <- rbind(
  part_function.total(6),
  part_function.total(7),
  part_function.total(8),
  part_function.total(9),
  part_function.total(10),
  part_function.total(11),
  part_function.total(12),
  part_function.total(13),
  part_function.total(14),
  part_function.total(15),
  part_function.total(16),
  part_function.total(17),
  part_function.total(18)
)

## Counting by age, and then head start participation

part_function <- function(age){
  part.df <<- subset(df,childage1986==age |childage1988==age |childage1990==age |childage1992==age |childage1994==age |
                      childage1996==age |childage1998==age |childage2000==age |childage2002==age |childage2004==age)$headstart
  table(part.df)
}

part.headstart <- rbind(
  part_function(6),
  part_function(7),
  part_function(8),
  part_function(9),
  part_function(10),
  part_function(11),
  part_function(12),
  part_function(13),
  part_function(14),
  part_function(15),
  part_function(16),
  part_function(17),
  part_function(18)
)

part.grade <- c(
  "Kindergarden",
  "1st Grade",
  "2nd Grade",
  "3rd Grade",
  "4th Grade",
  "5th Grade",
  "6th Grade",
  "7th Grade",
  "8th Grade",
  "9th Grade",
  "10th Grade",
  "11th Grade",
  "12th Grade"
)

part.school <- c(
  "Elementary School",
  "Elementary School",
  "Elementary School",
  "Elementary School",
  "Elementary School",
  "Elementary School",
  "Middle School",
  "Middle School",
  "Middle School",
  "High School",
  "High School",
  "High School",
  "High School"
)

## Creating Tables
  
part.table <- as.data.frame(cbind(part.age,part.grade,part.school,part.total,part.number,part.headstart))

names(part.table) <- c("Age","Grade","School Type","Total","PIAT Participants","Non-HeadStart","HeadStart")

table_part <- kable(part.table,caption="Children Summary")

# Linear Models as the Baseline

## Primary to elementary

df$compscore6to11 <- rowMeans(df[,c("compscoreage6","compscoreage7","compscoreage8","compscoreage9","compscoreage10","compscoreage11")],
                              na.rm=TRUE)

lm6to11 <- glm(
  data=df,
  formula=compscore6to11~
    Hispanic+Black+Male+
    FirstBorn+LogInc_0to3+MothED+Father_HH_0to3+PPVTat3+
    logBW+Repeat+
    PoorHealth+Age_Moth_Birth+
    headstart+headstart*Hispanic+headstart*Black+mentaldisability+
    learndisability)

## Middle School

df$compscore12to14 <- rowMeans(df[,c("compscoreage12","compscoreage13","compscoreage14")],na.rm=TRUE)

lm12to14 <- glm(
  data=df,
  formula=compscore12to14~
    Hispanic+Black+Male+
    FirstBorn+LogInc_0to3+MothED+Father_HH_0to3+PPVTat3+
    logBW+Repeat+
    PoorHealth+Age_Moth_Birth+
    headstart+headstart*Hispanic+headstart*Black+mentaldisability+
    learndisability)


## High School

df$compscore15to18 <- rowMeans(df[,c("compscoreage15","compscoreage16","compscoreage17","compscoreage18")],na.rm=TRUE)

#### Note that Father_HH_0to3 and PPVTat3 were removed becuase there were no values found at this level
lm15to18 <- glm(
  data=df,
  formula=compscore15to18~
    Hispanic+Black+Male+
    FirstBorn+LogInc_0to3+MothED+
    logBW+Repeat+
    PoorHealth+headstart*Hispanic+headstart*Black+Age_Moth_Birth+
    headstart+mentaldisability+learndisability)

# Table for linear models

table_lm <- stargazer::stargazer(lm6to11,lm12to14,lm15to18, type = 'latex',title="Composite Score based on Age",align=TRUE)

```
## Summary Table
We first begin with a summary table of the total number of children broken down by age, education level, and primary variables of interest.
```{r summary table,echo=FALSE}
table_part
```

As we have children born to females from the NLSY79 survey, there is a spread of children who become eligible for the Head Start Program at different years. We resolve this issue by tabulating the data to reflect specific criteria based on a child's age, rather than the current calendar year. Reading the table from the first row, we see that there are a total of 4,727 children in the survey who were aged 6 at some point in the survey time frame: 3,425 took a PIAT test at age 6, 3,939 did not participate in a Head Start program, and 788 participated in a Head Start program. Note, that the subset of 6 year old contain children who were born in 1979 up to 1998. This would give reason as to why for the last row of the table, there are only 3,295 children who were aged 18 - these are the children who were born from 1979 to 1986, and then through the course of the survey, continued to participate. 

## Linear Modeling of Composite Scores
Our first approach was to define an measure the impact of Head Start programs to the Peabody Picture Vocabulary Test, Peabody Individual Achievement Math Test, and Peabody Individual Achievement Reading Comprehension Test for each child. As the survey was administered annually up to 1994 and then biennially to 2004, we have target data of most children for each year, specifically whether or not they took the PIAT tests. But as the level of participation dropped as the survey moved forward, the data we have for effects when children are older are much smaller. The reasoning for this phenomena is currently unknown. Given this, a decision was made to categorize and analyze children based on age ranges rather than specific ages. This would allow an additional level of functional analysis on the school type level, namely, the effects of Head Start on Elementary, Middle, and High School children. 

```{r linear models, results = 'asis', echo=FALSE}
stargazer(lm6to11,lm12to14,lm15to18, type = 'latex',title="Composite Score based on Age",header=FALSE,single.row = TRUE)
```

In the dependent variables, we have broadly categorized individuals aged 6 to 11 as Elementary School children, 12 to 14 as Middle School, and 15 to 18 as High School; with **compscore6to11**, **compscore12to14**, and **compscore15to18**, reflecting each category respectively. While in the independent variables, we have a similar list of variables from Elementary School to Middle School. The reason why **Father_HH_0to3** and **PPVTat3** were excluded from the High School Linear Model was because at this point in time, there were no values reported in either variable for each children who participated in the PIAT testing - all values for these two variables were Not-Available. Because of the time frame of the survey, not all children consistently participated in the PIAT tests used to run our regression models.

In terms of magnitude (the absolute value of the coefficient) and significance (p-value), the independent variables that we see that are the biggest factors in a child's testing performance were:

* Race:
    + **Hispanic**: whether the child was Hispanic, 0 or 1.
    + **Black**: whether the child was Black, 0 or 1.
* Socioeconomic factors: 
    + **FirstBorn**: whether or not the child was the eldest among siblings, 0 or 1.
    + **LogInc_0to3**: the log scale income of the child from ages 0 to 3.
    + **Mother ED**: the mother's age in years/
    + **logBW**: a log scale of the child's weight at birth.
    + **Repeat**: whether or not the child ever repeated a grade, 0 or 1.
* Individual Factors
    + **mentaldisability**: whether or not this child was diagnosed with a mental disability, 0 or 1.
    + **learndisability**: whether or not this child was diagnosed with a learning disability, 0 or 1.
    
Overall, we find that the role of Head Start on a child's short term (Elementary School), to long term (High School), effects to be insignificant at best and negative in magnitude at worst. Whether or not this was attributed to the significant (almost 1/2 decrease) drop in participating children in the survey or because the effects of Head Start (in terms of what it teaches a child and its relevance to subsequent testing parameters) were not related to what was being assessed in the PIAT tests. What we can definitively was that throughout a child's formative years, mental health and learning disability, followed by whether a child has needed to repeat a grade level, and race, play a significant role in their success. When we interact race with participation with Head Start, we find that there are no significant effects found.

# Conclusion

# Bibliography

Anderson, Michael L. 2008. *"Multiple Inference and Gender Differences in the Effects of Early Intervention: A Reevaluation of the Abecedarian, Perry Preschool, and Early Training Projects."* Journal of the American Statistical Association, 103(484): 1481-95

Barnett, W. Steven. 1992. *"Benefits of Compensatory Preschool Education."* Journal of Human Resources, 27(2): 279-312

Barnett, W. Steven, Jason T. Hustedt, Allison H. Friedman, Judi Stevenson Boyd, and Pat Ainsworth. 2007. *The State of Preschool 2007: State Preschool Yearbook.* New Brunswick, NJ: National Institute for Early Education Research.

Belfield, Clive R., Milagros Nores, Steven Barnett, and Lawrence Schweinhart. 2006. *"The High/ Scope Perry Preschool Program: Cost-Benefit Analysis Using Data from the Age-40 Followup."* Journal of Human Resources, 40(1): 162-90.

Campbell, Frances A., Craig T. Ramey, Elizabeth Pungello, Joseph Sparling, and Shari Miller-Johnson. 2002. *"Early Childhood Education: Young Adult Outcomes From the Abecedarian Project."* Applied Developmental Science, 6(1): 42-57.

Case, Anne, Darren Lubotsky, and Christina Paxson. 2002. *"Economic Status and Health in Childhood: The Origins of the Gradient."* American Economic Review, 92(5): 1308-34.

Currie, Janet. 2001. *"Early Childhood Education Programs."* Journal of Economic Perspectives, 15(2): 213-38.

Currie, Janet, and Matthew Neidell. 2007. *"Getting inside the 'Black Box' of Head Start Quality: What Matters and What Doesn't."* Economics of Education Review, 26(1): 83-99.

Currie, Janet, and Duncan Thomas. 1995. *"Does Head Start Make a Difference?"* American Economic Review, 85(3): 341-64

Currie, Janet, and Duncan Thomas. 2000. *"School Quality and the Longer-Term Effects of Head Start."* Journal of Human Resources, 35(4): 755-74.

Garces, Eliana, Duncan Thomas, and Janet Currie. 2002. *"Longer-Term Effects of Head Start."* American Economic Review, 92(4): 999-1012.

Gormley, William T., Jr., and Ted Gayer. 2005. *"Promoting School Readiness in Oklahoma: An Evaluation of Tulsa's Pre-K Program."* Journal of Human Resources, 40(3): 533-58.

Heckman, James J. 2006. *"Skill Formation and the Economics of Investing in Disadvantaged Children."* Science, 312(5782): 1900-1902

Levin, Henry M., Clive Belfield, Peter Muennig, and Cecilia Rouse. 2007. *"The Public Returns to Public Educational Investments in African-American Males."* Economics of Education Review, 26(6): 699-708.

Ludwig, Jens, and Douglas L. Miller. 2007. *"Does Head Start Improve Children's Life Chances? Evidence from a Regression Discontinuity Design."* Quarterly Journal of Economics, 122(1): 159-208.

Ludwig, Jens, and Deborah A. Phillips. 2007. *"The Benefits and Costs of Head Start."* National Bureau of Economic Research Working Paper 12973

Ludwig, Jens, and Deborah A. Phillips. 2008. *"Long-Term Effects of Head Start on Low-Income Children."* Annals of the New York Academy of Sciences, 1136: 257-68.

Neal, Derek A., and William R. Johnson. 1996. *"The Role of Premarket Factors in Black-White Wage Differences."* Journal of Political Economy, 104(5): 869-95.
