---
title: 'Final Project: Headstart'
author: "Frank Chou, Tejaswi Pukkalla"
date: "April 22, 2019"
output: pdf_document
---

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(foreign)
df <- read.dta("nlsy_deming.dta")
```




# Abstract
This is frank's copy. 

# Results

## Table Means

```{r table means,echo=FALSE}


```

In this table, we have hte following understana;snaliwnfwafiwahfwakjf
In this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjf

This is normal text. However, when I reference a **variable** and when I have inserted code 

> This would be displayed as a block quote, it is great to display text or long quotes. snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this tasnaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this tasnaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this tasnaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this table, we have hte following understana;snaliwnfwafiwahfwakjfIn this ta

now I move on.