---
title: "Data Mining and Statistical Learning: Exercise 1"
author: "Frank Chou"
date: "February 9, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r, include=FALSE}
library(tidyverse)
library(reshape2)
library(expss)

df <- read.csv('greenbuildings.csv')

# Replicating the Analyst's results
df_rm10 <- df[df$leasing_rate > 10,] #this removes rows where leasing_rate is < 10
df_rm10_green <- df[df$green_rating == 1, ]
df_rm10_ngreen <- df[df$green_rating != 1, ]

rent <- median (df_rm10$Rent) #rent of all buildings
g_rent <- median (df_rm10_green$Rent) #rent of >10 occupancy green buildings
ng_rent <- median (df_rm10_ngreen$Rent) #rent of >10 occupancy non-green buildings
d_rent <- g_rent-ng_rent #difference in rent from green and non-green

n <- 30 #years of operation
o_rate <- 0.90 #occupancy rate
sqft <- 250000 #square footage of building
c_cost <- 100000000 #cost of building
g_cert <- 0.05 #percentage to certify green

#payoff analysis
p1 <- (c_cost*g_cert)/(d_rent*sqft) #at 100% occupancy
p2 <- (c_cost*g_cert)/(d_rent*sqft*o_rate) #at 90% occupancy

t1 <- (n-p1)*d_rent*sqft #total payoff under senario 1
t2 <- (n-p2)*d_rent*sqft*o_rate #total payoff under senario 2

full_rate <- c(100:1)/100 #a range of values from 0-100 for occupancy rates
full_payoff <- (c_cost*g_cert)/(d_rent*sqft*full_rate) #pay off times based on occupancy rates

df_plot <- data.frame(full_payoff,full_rate)

analyst_plot1 <- ggplot(df_plot)+
  geom_point(mapping = aes(x = full_payoff, y = full_rate), color='black')+
  labs(title="Payoff Analysis",x="Years to Payoff",y="Occupancy Rate")
```

# The Analyst's Analysis

```{r analyst plot 1, echo=FALSE}
analyst_plot1
```

  A rudimentary analysis of Analyst's parameters gives us a basic understanding of how long it would take to pay off the initial investment of building a green building: from a low of 7.6 years at 100% occupancy rate to over 800 years if we have a 1% occupancy rate. However, this is unrealistic, becuase if we were analyze the entire data set, we find that the rate of occupancy for green to non-green buildings are significantly different. But before we continue, let us focus on the actual building operational lifecycle of 30 years.

```{r analyst plot 2, echo=FALSE}
analyst_plot2 <-ggplot(df_plot)+
  geom_point(mapping = aes(x = full_payoff, y = full_rate), color='black')+
  labs(title="Payoff Analysis",subtitle="30 Year Operational Life",x="Years to Payoff",y="Occupancy Rate")+
  coord_cartesian(xlim=c(0, 30))+
  geom_hline(yintercept=1, linetype="dashed", color = "red")+
  geom_vline(xintercept=7.6, linetype="dashed", color = "red")+
  geom_hline(yintercept=0.25, linetype="dashed", color = "red")+
  geom_vline(xintercept=30, linetype="dashed", color = "red")
analyst_plot2
```

  Here we see a more accurate estimate of the payoff timetable for the building if we built it green. If we have 100% occupancy, we expect to pay it off within 7.6 years, however if we see 25% occupacy rates, we would break-even at the end of the building's estimated operational lifecycle, and any lower than 25% we would never break-even.

  Regardless of the analysis there are a number of points of fault within the Analyst's analysis. 

* One: by eliminating buildings where its occupancy rates are below 10%, we exclude important data that would give us insight as to how the market is doing, namely the fact that we don't expect occupancy rates to remain constant nor at max-capacity at the onsent of the building's commission. Becuase the data does not give us longitudial information regarding a given building's occupancy rate over the time of its lifetime, we do not have an idea of how a building going from non-green to green affects its rental value. 

* Two: by utilizing the median rental value of the subsets: green and non-green, we skew the data because we have more green versus non-green buildings:

```{r echo=FALSE}
point2 <- table(df$green_rating)
point2
```

Where "0" represent the number of non-green buildings and "1" represent the number of green buildings

* Three: As for the decision to utilize the median instead of the mean value of rent, a cursory assessment of different values provide different results of a green vs. non-green building.

```{r echo=FALSE}
hist(df_rm10$Rent,xlab="Rent",main ="Histogram of Rental Prices")
``` 


  A histogram of the Analyst's data set provides a view of how the rental prices are positively skewed. There are more buildings with rental prices in the 0-50 range than >50. Using the median value prices would have a result less than the mean value.
  
```{r echo=FALSE}
table <- matrix(c(
  median(df_rm10$Rent),
  mean(df_rm10$Rent)
    ),ncol=2,byrow=TRUE)

colnames(table) <- c("Median","Mean")
rownames(table) <- c("Whole Dataset")

table <- as.table(table)
table
```

  However, if we were to compare the price differences within the green versus non-green building subsets, we see that the effect of green buildings are less than anticipated. Intead of a 2.6 difference using median-values, it will only be 1.7. This will greatly extend our anticipated break-even point.
  
```{r echo=FALSE}
table <- matrix(c(
  median(df_rm10$Rent),
  mean(df_rm10$Rent),
  median(df_rm10_green$Rent),
  mean(df_rm10_green$Rent),
  median(df_rm10_ngreen$Rent),
  mean(df_rm10_ngreen$Rent)
),ncol=2,byrow=TRUE)

colnames(table) <- c("Median","Mean")
rownames(table) <- c("Whole Dataset","Green","Non-Green")

table <- as.table(table)
table
```

\newpage

# A Better Approach

  If we were to conduct the analysis with the full set of data and the mean-value, we see a different picture. 
  
```{r echo=FALSE}
df <- read.csv('greenbuildings.csv')
df_green <- df[df$green_rating == 1, ]
df_ngreen <- df[df$green_rating != 1, ]

rent <- mean (df$Rent) #rent of all buildings
g_rent <- mean (df_green$Rent) #rent of occupancy green buildings
ng_rent <- mean (df_ngreen$Rent) #rent of occupancy non-green buildings
d_rent <- g_rent-ng_rent #difference in rent from green and non-green

n <- 30 #years of operation
o_rate <- 0.90 #occupancy rate
sqft <- 250000 #square footage of building
c_cost <- 100000000 #cost of building
g_cert <- 0.05 #percentage to certify green

#payoff analysis
p1 <- (c_cost*g_cert)/(d_rent*sqft) #at 100% occupancy
p2 <- (c_cost*g_cert)/(d_rent*sqft*o_rate) #at 90% occupancy

t1 <- (n-p1)*d_rent*sqft #total payoff under senario 1
t2 <- (n-p2)*d_rent*sqft*o_rate #total payoff under senario 2

full_rate <- c(100:1)/100 #a range of values from 0-100 for occupancy rates
full_payoff <- (c_cost*g_cert)/(d_rent*sqft*full_rate) #pay off times based on occupancy rates

df_plot <- data.frame(full_payoff,full_rate)

# Begin Analysis of the better data

analyst_plot2 <-ggplot(df_plot)+
  geom_point(mapping = aes(x = full_payoff, y = full_rate), color='black')+
  labs(title="Payoff Analysis",subtitle="30 Year Operational Life",x="Years to Payoff",y="Occupancy Rate")+
  coord_cartesian(xlim=c(0, 30))+
  geom_hline(yintercept=1, linetype="dashed", color = "red")+
  geom_vline(xintercept=p1, linetype="dashed", color = "red")+
  geom_hline(yintercept=0.37, linetype="dashed", color = "red")+
  geom_vline(xintercept=30, linetype="dashed", color = "red")
analyst_plot2
```

  Here we see a more accurate estimate of the payoff timetable for the building if we built it green using mean-values. If we have 100% occupancy, we expect to pay it off within 11.43 years, however if we see 37% occupacy rates, we would break-even at the end of the building's estimated operational lifecycle, and any lower than 37% we would never break-even.

  In conclusion, we see that a more accurate assessment of the viability of building green is represented by using the full data set with mean-values instead of a truncated data set with median-values.

```{r,include=FALSE}
rm(list=ls())
```

\newpage

## Question 2

```{r load the data, include=FALSE}
library(plyr)
library(ggplot2)
library(maptools)
library(maps)
library(mapproj)
library(usmap)


abia <- read.csv('ABIA.csv')
airports <- read.csv('airports.csv')

# Let's visualize

abia <- select(abia,Origin,Dest)
airports <- select(airports,iata_code,name,latitude_deg,longitude_deg)

abia <- count(abia, vars = c("Origin", "Dest"))

origin_xy <- merge(abia, airports, by.x="Origin", by.y="iata_code")
names(origin_xy) <- c("origin", "destination","trips", "o_name", "oX", "oY")

dest_xy <-  merge(origin_xy, airports, by.x="destination", by.y="iata_code")
names(dest_xy) <- c("origin", "destination","trips", "o_name", "oY", "oX","d_name", "dY", "dX")

dest_xy <- dest_xy[- grep("Erase", dest_xy$o_name),]#erase the damn island of nowhere
dest_xy <- dest_xy[- grep("Erase", dest_xy$d_name),]#erase the damn island of nowhere

xquiet<- scale_x_continuous("", breaks=NULL)
yquiet<-scale_y_continuous("", breaks=NULL)
quiet<-list(xquiet, yquiet)

usa <- map_data("usa") # we already did this, but we can do it again
state <- map_data("state")

plot_1 <- ggplot()+
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = NA, color = "black")+
  geom_polygon(data = state,aes(x = long, y = lat, fill = region, group = group), color = "black")+
  guides(fill=FALSE)+  # do this to leave off the color legend
  geom_segment(data=dest_xy,aes(x=oX, y=oY,xend=dX, yend=dY, alpha=trips), col="black")+
  scale_alpha_continuous(range = c(0.09, 0.9))+
  theme(panel.background = element_rect(fill='white',colour='white'))+
  quiet+
  coord_fixed(1.3)+
  geom_point(data=dest_xy,aes(x=oX, y=oY) ,color="blue", size=1)+
  geom_point(data=dest_xy,aes(x=dX, y=dY) ,color="red", size=1)+
  ggtitle("Flights to and from Austin International Airport in 2008")+
  theme(plot.title = element_text(hjust = 0.5))
```
# A flight to Austin
  We can get a better understanding of which cities are connected to Austin by air. By plotting all of the flights to and from Austin to other cities in the United States, we can see which cities fly more to and from Austin. 
```{r plot_1,echo=FALSE}
plot_1
```

```{r,include=FALSE}
rm(list=ls())
```
  
## Question 3

